---
title: "Analyzing the Dynamics of Automobile Theft in Toronto"
subtitle: "A Decade of Surge in Comparison to Other Crimes, with a Focus on Persistent Spatial Disparities"
author: 
  - Sima Shmuylovich
thanks: "Code and data are available at: https://github.com/sshmuylovich/toronto-automobile-theft.git."
date: 25 January 2024
date-format: long
abstract: This paper investigates the changing landscape of automobile theft in Toronto over the past decade, comparing its growth to that of other crimes. Notably, this paper explores the intriguing phenomenon that despite the significant increase in automobile theft's prevalence, it has not expanded to encompass new neighborhoods. This paper aims to provide an in-depth analysis of automobile theft in Toronto, its evolution over the past decade, and its interplay with other crime categories as well as neighbourhood geography. In a broader context, this research contributes to the academic discourse on urban criminology and offers valuable insights for policymakers, law enforcement agencies, and urban planners.
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#### Preamble ####
# Purpose: Output a PDF that tells a story about Automobile Theft in Toronto using graphs and tables for support. 
# Author: Sima Shmuylovich
# Date: 25 January 2024
# Contact: sima.shmuylovich@mail.utoronto.ca
# License: MIT
# Pre-requisites: Run "import_packages.R", "01-download_data.R", and
# "02-data_cleaning.R"
```

```{r}
#| include: false
#| warning: false
#| message: false

library("tidyverse") 
library("knitr") 
library("opendatatoronto")
library("janitor") 
library("dplyr")
library("ggplot2") 
library("sf") 
library("patchwork")
```

# Introduction {#sec-introduction}
**Introduction**

Toronto, Canada's largest city, has witnessed various transformations in its urban landscape over the past decade. With these changes, the city's crime dynamics, particularly pertaining to automobile theft, have evolved. This paper presents a comprehensive analysis of automobile theft in Toronto from 2014 to 2023, offering a comparison with the progression of other major crimes during the same period. These include assault, bike theft, break and enter, robbery, shooting, homicide, and theft from a motor vehicle. This study also explores the geographical distribution of automobile theft across Toronto's 158 neighborhoods, providing insights into spatial trends and patterns [@citecrime]. 

The findings of this study are significant for several reasons. Firstly, they provide a nuanced understanding of how automobile theft, as a specific crime category, fits into the broader context of urban crime in Toronto. Secondly, the spatial analysis of crime distribution aids in identifying neighbourhoods that require focused policing and community intervention. Finally, this research contributes to the academic discourse on urban criminology, offering valuable insights for policymakers, law enforcement agencies, and urban planners. 

Two months ago, CP24 reported that auto theft reached 'historic highs' for Ontario in 2022 which saw Auto Insurers paying over \$1 billion in theft claims [@citecp24]. In May 2023, Ontario's solicitor general announced a budget of $51 million for the creation of a new organized crime and auto theft task force led by the Ontario Provincial Police [@citecbc]. It is all that more prevalent, therefore, to be conducting and publishing research that might aid in the slow-down of auto theft in Toronto.

The paper is structured as follows: following this introduction, the second section provides a review of relevant literature. The third section details the methodology, followed by a presentation and discussion of the findings in the fourth section. The paper concludes with implications for policy and suggestions for future research.

# Data {#sec-data}
The data used by the paper is 'Neighbourhood Crime Rates' [@citecrime], which was sourced from Open Data Toronto Portal. This dataset was accessed using the `opendatatoronto` library [@citeopendatatoronto]. The dataset was then simulated, cleaned, analyzed, and tested using the statistical programming language `R` [@citer], 'tidyverse' [@citetidyverse], 'janitor'[@citejanitor], 'dplyr' [@citedplyr], 'ggplot2' [@citeggplot2], 'sf' [@citesf], and 'knitr' [@citeknitr]. 


# Results {#sec-results}
```{r}
#| include: false
#| warning: false
#| message: false

# Read in Automobile Thefts Data
automobile_thefts_data <- read_csv("../data/cleaned_automobile_thefts_data.csv")

# Read in Map Data
map_data <- readRDS("../data/cleaned_map_data.rds") 
map_data_csv <- read_csv("../data/cleaned_map_data.csv") %>%
  filter(!is.na(num_automobile_thefts))

# Read in All Crime Thefts Data
all_crime_data <- read_csv("../data/cleaned_all_crime_data.csv")
```

### Automobile Theft Across All of Toronto Over the Past Decade (2014-2023) 
With auto theft having reached 'historic highs' for Ontario in 2022, this paper investigated how many vehicles were stolen in Toronto, Ontario's largest city, annually over the past decade. The following data was obtained:

```{r}
#| label: tbl-autothefttable
#| tbl-cap: The number of motor vehicles reported stolen each year from 2014 to 2023.
#| echo: false
#| warning: false
#| message: false

automobile_thefts_data %>% rename (
    "Year" = year,
    "Number of Automobile Thefts" = num_automobile_thefts
  ) %>%
  head() %>%
  kable() 
```

```{r}
#| include: false
#| warning: false
#| message: false
global_min <- min(map_data_csv$num_automobile_thefts)
global_max<- max(map_data_csv$num_automobile_thefts)

create_plot <- function(attribute, year) {
  ggplot() +
    geom_sf(data = map_data, aes_string(fill = attribute)) +
    scale_fill_gradient(low = "lightgrey", high = "purple", 
                        limits = c(global_min, global_max),
                        name = "Number of Auto Thefts") +
    labs(title = year) +
    theme_minimal() + 
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
}

plot1 <- create_plot("AUTOTHE25", "2014")
plot2 <- create_plot("AUTOTHE26", "2015")
plot3 <- create_plot("AUTOTHE27", "2016")
plot4 <- create_plot("AUTOTHE28", "2017")
plot5 <- create_plot("AUTOTHE29", "2018")
plot6 <- create_plot("AUTOTHE30", "2019")
plot7 <- create_plot("AUTOTHE31", "2020")
plot8 <- create_plot("AUTOTHE32", "2021")
plot9 <- create_plot("AUTOTHE33", "2022")
plot10 <- create_plot("AUTOTHE34", "2023")
```

### Automobile Theft By Neighbourhood Over the Past Decade (2014-2023) 
With 158 neighbourhood's in Toronto, this paper investigated which neighbourhood has had the most difficulty with auto theft over the past decade. The following data was obtained:
```{r}
#| label: fig-autotheftmap1
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2014.
#| echo: false
#| warning: false
#| message: false
plot1
```

```{r}
#| label: fig-autotheftmap2
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2015.
#| echo: false
#| warning: false
#| message: false
plot2
```

```{r}
#| label: fig-autotheftmap3
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2016.
#| echo: false
#| warning: false
#| message: false
plot3
```

```{r}
#| label: fig-autotheftmap4
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2017.
#| echo: false
#| warning: false
#| message: false
plot4
```

```{r}
#| label: fig-autotheftmap5
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2018.
#| echo: false
#| warning: false
#| message: false
plot5
```

```{r}
#| label: fig-autotheftmap6
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2019.
#| echo: false
#| warning: false
#| message: false
plot6
```

```{r}
#| label: fig-autotheftmap7
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2020.
#| echo: false
#| warning: false
#| message: false
plot7
```

```{r}
#| label: fig-autotheftmap8
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2021.
#| echo: false
#| warning: false
#| message: false
plot8
```

```{r}
#| label: fig-autotheftmap9
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2022.
#| echo: false
#| warning: false
#| message: false
plot9
```

```{r}
#| label: fig-autotheftmap10
#| fig-cap: The number of motor vehicles reported stolen from each neighbourhood in 2023.
#| echo: false
#| warning: false
#| message: false
plot10
```
# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

## Second discussion point {#sec-second-point}

## Third discussion point {#sec-third-point}

## Weaknesses and next steps {#sec-weaknesses-and-next-steps}

\newpage

# References {#sec-references}

