---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#### Preamble ####
# Purpose: ....
# Author: Sima Shmuylovich
# Date: 25 January 2024
# Contact: sima.shmuylovich@mail.utoronto.ca
# License: MIT
# Pre-requisites: Run "import_packages.R", "01-download_data.R", and
# "02-data_cleaning.R"
```

```{r}
#| include: false
#| warning: false
#| message: false

library("tidyverse") 
library("knitr") 
library("opendatatoronto")
library("janitor") 
library("dplyr")
library("ggplot2") 
library("sf") 
library("patchwork")
```


# Introduction

# Data {#sec-data}

# Results
```{r}
#| include: false
#| warning: false
#| message: false

map_data <- readRDS("../data/cleaned_map_data.rds") 
map_data_csv <- read_csv("../data/cleaned_map_data.csv") %>%
  filter(!is.na(num_automobile_thefts))

global_min <- min(map_data_csv$num_automobile_thefts)
global_max<- max(map_data_csv$num_automobile_thefts)

create_plot <- function(attribute, year) {
  ggplot() +
    geom_sf(data = map_data, aes_string(fill = attribute)) +
    scale_fill_gradient(low = "lightgrey", high = "purple", 
                        limits = c(global_min, global_max),
                        name = "Automobile-Thefts") +
    labs(title = year) +
    theme_minimal() + 
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
}

plot1 <- create_plot("AUTOTHE25", "2014")
plot2 <- create_plot("AUTOTHE26", "2015")
plot3 <- create_plot("AUTOTHE27", "2016")
plot4 <- create_plot("AUTOTHE28", "2017")
plot5 <- create_plot("AUTOTHE29", "2018")
plot6 <- create_plot("AUTOTHE30", "2019")
plot7 <- create_plot("AUTOTHE31", "2020")
plot8 <- create_plot("AUTOTHE32", "2021")
plot9 <- create_plot("AUTOTHE33", "2022")
plot10 <- create_plot("AUTOTHE34", "2023")

plot_layout <- (plot1 + plot2 + plot3 + plot4 + plot5 + plot6 + plot7 + plot8 + plot9 + plot10)
```

```{r}
ggsave("combined_plots.png", plot_layout, width = 20, height = 10, units = "cm")
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point


## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}



\newpage


# References


